---
title: "DS CP"
author: "B1 Group 2"
output:
  html_document:
    df_print: paged
  html_notebook: default
  word_document: default
---

```{r}
# Load Libraries
library(class)
library(e1071)
library(caTools)
library(caret)
library(dplyr)
library(pROC)
library(ggplot2)
library(kknn)
library(plotly)
library(sunburstR)
```


Import Dataset
```{r}
emp_data <- read.csv("C:/Users/bhagw/OneDrive/Desktop/employee_attrition.csv")
head(emp_data)
```

Exclude Non-Informative Columns
```{r}
emp_data <- emp_data %>% select(-EmployeeCount, -Over18, -StandardHours)
```

Convert Target Variable to Factor
```{r}
emp_data$Attrition <- as.factor(emp_data$Attrition)
```


Visualization: Distribution of Attrition
```{r}
attrition_counts <- table(emp_data$Attrition)
labels <- c("No Attrition", "Yes Attrition")
fig <- plot_ly(
  labels = labels,
  values = attrition_counts,
  type = "pie",
  textfont = list(size = 15),
  marker = list(
    colors = c('#B9C0C9', 'yellow'),
    line = list(color = '#000000', width = 1.5)
  )
) %>%
  layout(title = "Distribution of Attrition Variable")
fig
```


Visualization: Effect of Age on Attrition
```{r}
ggplot(emp_data, aes(x = Age, fill = as.factor(Attrition))) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 30) +
  scale_fill_manual(values = c("#B8890B", "#FFB90F"), labels = c("No Attrition", "Yes Attrition")) +
  labs(x = "Age", y = "Count", fill = "Attrition") +
  ggtitle("Distribution of Age by Attrition") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_density(aes(y = ..count..), color = "black", adjust = 1)
```

Visualization: Gender Distribution
```{r}
left_counts <- emp_data %>%
  group_by(Gender, Attrition) %>%
  summarise(count = n(), .groups = 'drop')
stayed_data <- left_counts %>% filter(Attrition == "No")
left_data <- left_counts %>% filter(Attrition == "Yes")
```


Stayed Pie Chart
```{r}
fig1 <- plot_ly(
  labels = paste(stayed_data$Gender, "Stayed"),
  values = stayed_data$count,
  type = "pie",
  marker = list(colors = c('#FF9999', '#66B3FF'), line = list(color = '#000000', width = 1.5)),
  textinfo = 'percent+label'
) %>%
  layout(title = "Gender Distribution for Employees Who Stayed")
fig1
```


Left Pie Chart
```{r}
ggplot(left_data, aes(x = "", y = count, fill = Gender)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar(theta = "y") +
  labs(title = "Gender Distribution for Employees Who Left") +
  theme_void() +
  scale_fill_manual(values = c("lightblue", "lightpink"))
```

Data Splitting
```{r}
set.seed(123)
split <- sample.split(emp_data$Attrition, SplitRatio = 0.7)
train_data <- subset(emp_data, split == TRUE)
test_data <- subset(emp_data, split == FALSE)
```


Data Preparation for k-NN
```{r}
train_x <- train_data[, sapply(train_data, is.numeric)]
test_x <- test_data[, sapply(test_data, is.numeric)]
train_y <- train_data$Attrition
test_y <- test_data$Attrition
```


Normalize Numeric Data
```{r}
normalize <- function(x) { (x - min(x)) / (max(x) - min(x)) }
train_x_norm <- as.data.frame(lapply(train_x, normalize))
test_x_norm <- as.data.frame(lapply(test_x, normalize))
```


KNN MODEL
```{r}
knn_prediction <- knn(train = train_x_norm, test = test_x_norm, cl = train_y, k = 5)
knn_confMatrix <- confusionMatrix(knn_prediction, test_y)
print(knn_confMatrix)
```

KNN ROC CURVE
```{r}
knn_model_kknn <- train.kknn(Attrition ~ ., data = train_data, kmax = 5, distance = 2)
knn_probabilities <- predict(knn_model_kknn, test_data, type = "prob")
knnroc_obj <- roc(test_y, knn_probabilities[, 2])  # Probabilities for "Yes" class
ggroc(knnroc_obj) +
  ggtitle(paste("k-NN ROC Curve (AUC =", round(auc(knnroc_obj), 2), ")")) +
  xlab("1 - Specificity (False Positive Rate)") +
  ylab("Sensitivity (True Positive Rate)") +
  theme_minimal()
```

SVM 
```{r}
svm_model <- svm(Attrition ~ ., data = train_data, kernel = "radial", probability = TRUE)
svm_predictions <- predict(svm_model, newdata = test_data, probability = TRUE)
svm_confMatrix <- confusionMatrix(svm_predictions, test_data$Attrition)
print(svm_confMatrix)
```
SVM ROC CURVE
```{r}
svm_probabilities <- attr(svm_predictions, "probabilities")[, 2] 
svmroc_obj <- roc(test_y, svm_probabilities)
ggroc(svmroc_obj) +
  ggtitle(paste("SVM ROC Curve (AUC =", round(auc(svmroc_obj), 2), ")")) +
  xlab("1 - Specificity (False Positive Rate)") +
  ylab("Sensitivity (True Positive Rate)") +
  theme_minimal()
```

LOGISTIC REGRESSION
```{r}
log_model <- glm(Attrition ~ ., family = binomial, data = train_data)
log_pred_prob <- predict(log_model, newdata = test_data, type = "response")
log_pred_class <- ifelse(log_pred_prob > 0.5, "Yes", "No")
log_pred_class <- as.factor(log_pred_class)
log_confMatrix <- confusionMatrix(log_pred_class, test_data$Attrition)
print(log_confMatrix)
```
LOGISTIC ROC CURVE
```{r}
logroc_obj <- roc(as.numeric(test_data$Attrition) - 1, log_pred_prob)
ggroc(logroc_obj) +
  ggtitle(paste("logistic ROC Curve (AUC =", round(auc(logroc_obj), 2), ")")) +
  xlab("1 - Specificity (False Positive Rate)") +
  ylab("Sensitivity (True Positive Rate)") +
  theme_minimal()
```

